### 2025-08-05 paper/camera/sects/experiments.tex
compact more
ready to submit

In this section, we evaluate \fname{} against both PPO (for single agent settings), MAPPO (for multi-agent settings), and SHAC across various multi-agent scenarios and architectures. An ablation study replacing only the transition function (SHAC+, where the reward function remains differentiable) is provided in \Cref{apx:ablation}.
We select PPO as our baseline because, despite the emergence of more recent algorithms HAPPO \cite{Kuba2021TrustRM}, and SAC \cite{Haarnoja2018SoftAA}, it remains one of the few approaches that scales effectively to multi-agent settings without substantial modifications (e.g., MAPPO~\cite{DBLP:conf/nips/YuVVGWBW22}). This makes PPO a representative benchmark for state-of-the-art performance in non-differentiable environments.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{figs/grads-transformer-transport.pdf}
    \caption{Policy gradient norms in the Transport environment over 5000 epochs. Spikes in gradient norm for SHAC (epoch 2700) and \fname{} (epoch 3200) correspond to generalization events when agents begin frequent interactions with the package and each other.}\label{fig:grads-transformer-transport}
    \vspace{0.5cm}
\end{figure}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/main-transformer.pdf}
    \caption{Performance comparison of \fname{}, PPO/MAPPO, and SHAC across different scenarios (Dispersion, Transport, Discovery, and Sampling) using the Transformer architecture for multi-agent settings and the MLP architecture for single-agent settings. The results show the mean and standard deviation of rewards over 3 runs. For results using the MLP architecture in all scenarios, refer to \Cref{apx:fig:experiments-mlp}.}%
    \label{fig:experiments}
\end{figure*}

\begin{table}[t]
    \centering
    \input tables/max.tex
    \vspace{0.2cm}
    \caption{Normalized rewards (relative to the best performing model) for the different scenarios. Best results are in bold. \textsc{D} stands for Dispersion, \textsc{T} for Transport, \textsc{Di} for Discovery, and \textsc{S} for Sampling.}\label{tab:max-rewards}.
\end{table}

>>>>

In this section, we evaluate \fname{} against both PPO (for single agent settings), MAPPO (for multi-agent settings), and SHAC across various multi-agent scenarios and architectures. An ablation study replacing only the transition function (SHAC+, where the reward function remains differentiable) is provided in \Cref{apx:ablation}.
We select PPO as our baseline because, despite the emergence of more recent algorithms HAPPO \cite{Kuba2021TrustRM}, and SAC~\cite{DBLP:conf/icml/HaarnojaZAL18}, it remains one of the few approaches that scales effectively to multi-agent settings without substantial modifications (e.g., MAPPO~\cite{DBLP:conf/nips/YuVVGWBW22}). This makes PPO a representative benchmark for state-of-the-art performance in non-differentiable environments.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.65\columnwidth]{figs/grads-transformer-transport.pdf}
    \caption{Policy gradient norms in the Transport environment over 5000 epochs. Spikes in gradient norm for SHAC (epoch 2700) and \fname{} (epoch 3200) correspond to generalization events when agents begin frequent interactions with the package and each other.}\label{fig:grads-transformer-transport}
    \vspace{0.1cm}
\end{figure}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/main-transformer.pdf}
    \caption{Performance comparison of \fname{}, PPO/MAPPO, and SHAC across different scenarios (Dispersion, Transport, Discovery, and Sampling) using the Transformer architecture for multi-agent settings and the MLP architecture for single-agent settings. The results show the mean and standard deviation of rewards over 3 runs. For results using the MLP architecture in all scenarios, refer to \Cref{apx:fig:experiments-mlp}.}%
    \label{fig:experiments}
    \vspace{0.1cm}
\end{figure*}

\begin{table}[t]
    \centering
    \input tables/max.tex
    \vspace{0.1cm}
    \caption{Normalized rewards (relative to the best performing model) for the different scenarios. Best results are in bold. \textsc{D} stands for Dispersion, \textsc{T} for Transport, \textsc{Di} for Discovery, and \textsc{S} for Sampling.}\label{tab:max-rewards}.
\end{table}

### 2025-08-05 paper/camera/sects/limitations.tex
ready to submit

\paragraph{World Model Learning and Cold-Start Problem}
A central challenge lies in learning the world model, which depends on rich agent-environment interactions. Sparse or delayed rewards create a cold-start problem: until the world model is sufficiently trained, agents struggle to explore effectively. This can be addressed by seeding with a cold-start dataset---manually designed or generated using algorithms like PPO or MPPI~\cite{Alvarez24}---to bootstrap the world model and provide a stable training foundation.

@@

\paragraph{Computational Overhead}
\fname{} addresses SHAC's limitations by adding transition and reward networks, but this increases computational overhead. Training burden is especially pronounced with many agents, where the transition network must capture complex inter-agent dynamics. While this trade-off is often acceptable, it significantly slows training and demands substantial computational resources.

>>>>

\paragraph{World Model Learning and Cold-Start Problem}
A central challenge lies in learning the world model from agent-environment interactions. Sparse or delayed rewards create a cold-start problem: until sufficiently trained, agents struggle to explore effectively. This can be addressed by seeding with a cold-start dataset---manually designed or generated using algorithms like PPO or MPPI~\cite{Alvarez24}---to bootstrap the world model and provide stable training foundation.

@@

\paragraph{Computational Overhead}
\fname{} addresses SHAC's limitations by adding transition and reward networks, increasing computational overhead. Training burden is especially pronounced with many agents, where the transition network must capture complex inter-agent dynamics. While often acceptable, this significantly slows training and demands substantial resources.
